---
title: "RD Robustness Project"
author: "Matthew Hong & Maor Milgrom"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
params:
  exercise:
    value: zero
    choices: [zero, bwo]
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, fig.height=4.5, root.dir = '~/Documents/Boston/PhD/Research/2021_Spring_RA/RDSimulation/code')
setwd('~/Documents/Boston/PhD/Research/2021_Spring_RA/RDSimulation/code')  
```


# Introduction
The goal of this project is to test the robustness of the Regression-Discontinuity analysis to different extreme cases, via simulations, using the 'rdrobust' package.  

## This Document: Excercise \#2
Stressing out results, by adding noise inside Optimal Bandwitdh (OBW), before repeating exercise #1.

### Select exercise type:
```{r select_exercise}
exercise=params$exercise      # Type of exercise - "zero" or "bwo" ('bandwidth only')
```


```{r subtitle, include=F}
subtitle_var = ifelse(exercise=="zero", "replacing with zeros", "dropping outside OBW")
```

---
subtitle: "Exercise: stressout before `r subtitle_var`"
---


```{r packages, include=F}
library(dplyr)
library(rdrobust)
library(ggplot2)
library(reshape2)
library(knitr)
library(data.table)
set.seed(1)
```


### Set Parameters
Here we set the main parameters for the excercise:

```{r parameters}
jump=10           # Size of jump at cutoff
loop=1000
figs.iter.save=5
quadratic=T       # T - quadratic DGP, F - linear
symm_obw="mserd"  # mserd - symmetric OBW, msetwo - asymmetric OBW 
normal.x=T        # T - normal draws of x around cutoff, F - uniform draws
dgp.sd=10         # sd of normal noise added to DGP
noisy.sd=30       # sd of noise added inside OBW
so.int=2          # interval inside OBW for adding noise
bc=F              # bias-corrected estimates or conventional
```

```{r setup, include=F}
### dataframes for storing results
results<- as.data.table(matrix(0, ncol = 10, nrow = loop))
colnames(results) <- c("coef","obw","obw_l","obw_r",
                       "coef_so","obw_so","obw_l_so","obw_r_so",
                       "coef_diff","obw_diff")

results.exercise<- as.data.table(matrix(0, ncol = 8, nrow = loop))
colnames(results.exercise) <- c("coef","obw","obw_l","obw_r",
                            "coef_diff","obw_diff",
                            "coef_diff_so","obw_diff_so")

### Extensions for files based on parameters selected
figs.dir="/Users/MH/Documents/Boston/PhD/Research/2021_Spring_RA/RDSimulation/figures/"
quad_test=ifelse(quadratic==T,"quadratic","linear")
normal_test=ifelse(normal.x==T,"rnorm","runif")
bc_test=ifelse(bc==T,"bc","c")
save.ext=paste(bc_test,quad_test,normal_test,sep = "_")
file.var=paste("summary_",save.ext,".html",sep = "")
begin.figures=loop-figs.iter.save
```


### Simulate DGP
```{r dgp}
dt <- as.data.table(matrix(0, ncol = 0, nrow = length(seq(-100,100,0.01))))
dt$x <- round(seq(-100,100,0.01), digits=2)
dt <- dt[x!=0]
dt$treated <- ifelse(dt$x>0, 1, 0)

dt$y.model<- 0.5*dt$x  - 0.025*dt$x^2*quadratic + jump*dt$treated
dt$y <- dt$y.model + rnorm(length(dt$x),0,dgp.sd)
dt$y.noisy <- dt$y + rnorm(length(dt$x),0,noisy.sd)

## dataframe for draws (samples)
sample.x <- as.data.table(matrix(0, ncol = 0, nrow = nrow(dt)/10))
```

### Iterations
Running `r loop` iterations, and saving figures from `r figs.iter.save` last iterations to file.
In each iteration, we: \newline

1. draw randomly `r nrow(dt)/10` observations around the cutoff.
2. compute the OBW and RD coefficient
3. Add noise just inside OBW (within `r so.int` units)
4. compute again the OBW and RD coefficient
5. drop observations outside the new OBW/replace them with zero
6. compute again the OBW and RD coefficient.

```{r iterations, include=F}
for(i in 1:loop) {
  if (normal.x==T) {
    sample.x$x <- round(rnorm(nrow(dt)/10, 0, 10),digits = 2) 
  } else {
    sample.x$x <- round(runif(nrow(dt)/10, -20,20),digits = 2) 
  }

  sample.x <- sample.x[x %between% c(-100,100)]
  sample <- merge(dt,sample.x, by="x")
  
  results.current <- rdrobust(sample$y,sample$x, bwselect = symm_obw)
  results[i,"coef" := results.current$coef[bc+1]-jump] # normalizing to zero
  results[i,c("obw_l","obw_r") := as.list(results.current$bws[bc+1,])]
  results[i,"obw" := sum(obw_l,obw_r)]
  
  sample[, y.so := fifelse(x %between% c(results[i,4]-so.int, results[i,4]) |
                             x %between% c(-results[i,3], -results[i,3]+so.int),
                           sample$y.noisy,sample$y)]
  
  results.current <- rdrobust(sample$y.so,sample$x,bwselect = symm_obw)
  results[i,"coef_so" :=results.current$coef[bc+1]-jump] # normalizing to zero
  results[i,c("obw_l_so","obw_r_so") := as.list(results.current$bws[bc+1,1:2])]
  results[i,"obw_so" := sum(obw_l_so,obw_r_so)]
  results[i,c("coef_diff","obw_diff") := (results[i,c("coef","obw")]-results[i,c("coef_so","obw_so")])]
  
  if (exercise=="zero") {
    sample.exercise <- copy(sample)
    sample.exercise[, y.so := fifelse(sample$x %between% c(-results[i,"obw_l_so"], results[i,"obw_r_so"]), sample$y.so, 0)]
    
  } else if (exercise=="bwo") {
    sample.exercise <- sample[x %between% c(-results[i,"obw_l_so"],results[i,"obw_r_so"])]
  }
  
    results.current <- rdrobust(sample.exercise$y.so,sample.exercise$x,bwselect = symm_obw)
    results.exercise[i,"coef" :=results.current$coef[bc+1]-jump] # normalizing to zero
    results.exercise[i,c("obw_l","obw_r") := as.list(results.current$bws[bc+1,1:2])]
    results.exercise[i,"obw" := sum(obw_l, obw_r)]
    results.exercise[i,c(5,6) := results[i,c("coef","obw")]-results.exercise[i,1:2]]
    results.exercise[i,c(7,8) := results[i,c("coef_so","obw_so")]-results.exercise[i,1:2]]
    
### FIGURES INSIDE LOOP - SPECIFIC DRAWS ###
if (i >= begin.figures) {

  coef_base=paste("Base: Coef = ",round(results[i,1],digits = 2)+jump,sep = "")
  coef_so=paste("so: Coef = ",round(results[i,5],digits = 2)+jump,sep = "")
  coef_treat=paste(exercise,": Coef = ",round(results.exercise[i,1],digits = 2)+jump,sep = "")


  figure_name=paste(figs.dir,"sample_so",i,"_",exercise,"_",save.ext,".png",sep = "")
  png(figure_name)
  temp.exercise <- sample.exercise[x %between% c(-20,20)]
  temp <- sample[x %between% c(-20,20)]
  plot(temp$y.so~temp$x, ylim = range(c(temp.exercise$y.so,temp$y.so)),
       xlim = range(c(temp.exercise$x,temp$x)),
       pch=20,
       col="red", ylab = "Y", xlab = "X")
  par(new = T)
  plot(temp.exercise$y.so~temp.exercise$x, ylim = range(c(temp.exercise$y.so,temp$y.so)),
       xlim = range(c(temp.exercise$x,temp$x)),
       pch=20,
       col="blue", ylab = "Y", xlab = "X")
  abline(v = c(-results[i,3],-results[i,"obw_l_so"], -results.exercise[i,3],
               results[i,4], results[i,"obw_r_so"], results.exercise[i,4]),
         col=c("blue","red","green", "blue", "red","green"),
         lty=c(1,2,3,1,2,3), lwd=c(1,2,3,1,2,3))
  legend("topleft", legend=c(coef_base, coef_so, coef_treat),
         col=c("blue", "red", "green"), lty=1:3, cex=0.8)
  dev.off()
  
    figure_name=paste(figs.dir,"sample_so",i,"_",exercise,"_",save.ext,"_","model.png",sep = "")
    png(figure_name)
    temp.exercise <- sample.exercise[x %between% c(-20,20)]
    temp <- sample[x %between% c(-20,20)]
    plot(temp$y.so~temp$x, ylim = range(c(temp.exercise$y.so,temp$y.so)),
         xlim = range(c(temp.exercise$x,temp$x)),
         col="red", ylab = "Y", xlab = "X")
    par(new = T)
    plot(temp.exercise$y.so~temp.exercise$x, ylim = range(c(temp.exercise$y.so,temp$y.so)),
         xlim = range(c(temp.exercise$x,temp$x)),
         col="blue", ylab = "Y", xlab = "X")
    par(new = T)
    plot(temp$y.model~temp$x, ylim = range(c(temp.exercise$y.so,temp$y.so)),
         xlim = range(c(temp.exercise$x,temp$x)),
         col="black", ylab = "Y", xlab = "X")
    abline(v = c(-results[i,3],-results[i,"obw_l_so"], -results.exercise[i,3],
               results[i,4], results[i,"obw_r_so"], results.exercise[i,4]),
           col=c("blue","red","green", "blue", "red","green"),
           lty=c(1,2,3,1,2,3), lwd=c(1,2,3,1,2,3))
    legend("topleft", legend=c(coef_base, coef_so, coef_treat),
           col=c("blue", "red", "green"), lty=1:3, cex=0.8)
    dev.off()
    
    
    figure_name=paste(figs.dir,"rd_plot_so",i,"_",save.ext,".png",sep = "")
    png(figure_name)
    rdplot(sample$y.so,sample$x)
    dev.off()
    

    figure_name=paste(figs.dir,"sample",i,"_",save.ext,".png",sep = "")
    png(figure_name)
    sample[x %between% c(-30,30)] %T>%
      plot(y~x,., ylim = range(c(y,y.model)),
           col="blue", ylab = "Y", xlab = "X") %T>%
      par(new = T) %>%
      plot(y.model~x,., ylim = range(c(y,y.model)),
           axes = FALSE, xlab = "", ylab = "")
    dev.off()
    }
}

```

### Presenting figures from last iteration for illustration
```{r sample_figures_present, echo=F}
  i=loop

  coef_base=paste("Base: Coef = ",round(results[i,1],digits = 2)+jump,sep = "")
  coef_so=paste("so: Coef = ",round(results[i,5],digits = 2)+jump,sep = "")
  coef_treat=paste(exercise,": Coef = ",round(results.exercise[i,1],digits = 2)+jump,sep = "")

    
    rdplot(sample$y.so,sample$x, title = "RD Plot with stressout")

    temp.exercise <- sample.exercise[x %between% c(-20,20)]
    temp=sample[x %between% c(-20,20)]
    plot(temp$y.so~temp$x, ylim = range(c(temp.exercise$y.so,temp$y.so)),
         xlim = range(c(temp.exercise$x,temp$x)),
         col="red", ylab = "Y", xlab = "X")
    par(new = T)
    plot(temp.exercise$y.so~temp.exercise$x, ylim = range(c(temp.exercise$y.so,temp$y.so)),
         xlim = range(c(temp.exercise$x,temp$x)),
         col="blue", ylab = "Y", xlab = "X")
    par(new = T)
    plot(temp$y.model~temp$x, ylim = range(c(temp.exercise$y.so,temp$y.so)),
         xlim = range(c(temp.exercise$x,temp$x)),
         col="black", ylab = "Y", xlab = "X")
    abline(v = c(-results[i,3],-results[i,"obw_l_so"], -results.exercise[i,3],
               results[i,4], results[i,"obw_r_so"], results.exercise[i,4]),
           col=c("blue","red","green", "blue", "red","green"),
           lty=c(1,2,3,1,2,3), lwd=c(1,2,3,1,2,3))
    legend("topright", legend=c(coef_base, coef_so, coef_treat),
           col=c("blue", "red", "green"), lty=1:3, cex=0.8)
    title(main = paste0("Comparing baseline to streesout to ",exercise))


    rdplot(sample.exercise$y,sample.exercise$x, title = paste("RD Plot for",exercise,"with streesout"))

```

## Results

### Figures summarizing iterations
```{r figures_summary_print, echo=F, fig.height=3}
## PDF'S OF LEVELS
ggplot(results, aes(coef_so)) + stat_density(geom="line") + 
   geom_vline(aes(xintercept=0), colour="#BB0000", linetype="dashed") +
  ggtitle("Density of RD coefficients: stressout")

ggplot(results.exercise, aes(coef)) + stat_density(geom="line") +
  geom_vline(aes(xintercept=0), colour="#BB0000", linetype="dashed") +
    ggtitle(paste0("Density of RD coefficients: ",exercise))

  ## so_base_diff_CDF
 ggplot(results, aes(coef_diff)) + stat_ecdf(geom = "step") +
  ggtitle("CDF of difference b/w baseline coefficients and stressout")

 ggplot(results, aes(obw_diff)) + stat_ecdf(geom = "step") +
  ggtitle("CDF of difference b/w baseline OBW and stressout")

### treat_so_diff
  ggplot(results.exercise, aes(coef_diff_so)) + stat_ecdf(geom = "step") +
  ggtitle(paste0("CDF of difference b/w stressout coefficients and ",exercise))

  ggplot(results.exercise, aes(obw_diff_so)) + stat_ecdf(geom = "step") +
  ggtitle(paste0("CDF of difference b/w stressout OBW and ",exercise))

```

```{r figures_summary_save, include=F}
## PDF'S OF LEVELS
figure_name=paste(figs.dir,"coef","_",save.ext,".png",sep = "")
png(figure_name)
ggplot(results, aes(coef)) + stat_density(geom="line") + 
  geom_vline(aes(xintercept=0), colour="#BB0000", linetype="dashed")
dev.off()

figure_name=paste(figs.dir,"coef_so","_",save.ext,".png",sep = "")
png(figure_name)
ggplot(results, aes(coef_so)) + stat_density(geom="line") + 
   geom_vline(aes(xintercept=0), colour="#BB0000", linetype="dashed")
dev.off()

figure_name=paste(figs.dir,"coef","_",exercise,"_",save.ext,".png",sep = "")
png(figure_name)
ggplot(results.exercise, aes(coef)) + stat_density(geom="line") +
  geom_vline(aes(xintercept=0), colour="#BB0000", linetype="dashed")
dev.off()

  ## so_base_diff_CDF
figure_name=paste(figs.dir,"coef_diff_cdf_base_so","_",save.ext,".png",sep = "")
png(figure_name)
 ggplot(results, aes(coef_diff)) + stat_ecdf(geom = "step")
 dev.off()
 
 figure_name=paste(figs.dir,"obw_diff_cdf_base_so","_",save.ext,".png",sep = "")
 png(figure_name)
 ggplot(results, aes(obw_diff)) + stat_ecdf(geom = "step")
 dev.off()

### treat_base_diff
 figure_name=paste(figs.dir,"coef_diff_cdf_base","_",exercise,"_",save.ext,".png",sep = "")
 png(figure_name)
 ggplot(results.exercise, aes(coef_diff)) + stat_ecdf(geom = "step")
 dev.off()
 
 figure_name=paste(figs.dir,"obw_diff_cdf_base","_",exercise,"_",save.ext,".png",sep = "")
 png(figure_name)
 ggplot(results.exercise, aes(obw_diff)) + stat_ecdf(geom = "step")
 dev.off()

### treat_so_diff
  figure_name=paste(figs.dir,"coef_diff_cdf_so","_",exercise,"_",save.ext,".png",sep = "")
  png(figure_name)
  ggplot(results.exercise, aes(coef_diff_so)) + stat_ecdf(geom = "step")
  dev.off()

  figure_name=paste(figs.dir,"obw_diff_cdf_so","_",exercise,"_",save.ext,".png",sep = "")
  png(figure_name)
  ggplot(results.exercise, aes(obw_diff_so)) + stat_ecdf(geom = "step")
  dev.off()

```

### Summary results - Table 
Note: coefficient (treatment effects) are normalized to zero, by subtracting from each estimate the size of the jump at the cutoff.
```{r table, echo=F}
results_table=as.data.table(matrix(0, ncol = 5, nrow = 2))
results_table[1,1]=round(mean(results$coef),digits = 4)
results_table[1,2]=round(mean(results$coef_so),digits = 4)
results_table[1,3]=round(mean(results.exercise$coef),digits = 4)
results_table[1,4]=round(mean(results.exercise$coef_diff),digits = 4)
results_table[1,5]=round(mean(results.exercise$coef_diff_so),digits = 4)
results_table[2,1]=round(mean(results$obw),digits = 4)
results_table[2,2]=round(mean(results$obw_so),digits = 4)
results_table[2,3]=round(mean(results.exercise$obw),digits = 4)
results_table[2,4]=round(mean(results.exercise$obw_diff),digits = 4)
results_table[2,5]=round(mean(results.exercise$obw_diff_so),digits = 4)
colnames(results_table) = c("base","stressout",exercise,"diff_base","diff_so")
rownames(results_table) = c("coef","obw")

kable(results_table, caption = "Summary Table")
```

### Interpreting results
`r ifelse(exercise=="zero","Adding noise within OBW leads to new OBW to be narrower, but coefficients remain unbiased. Then, when replacing values outside new OBW with zeros, the estimated coefficients remain unbiased","Adding noise within OBW leads to new OBW to be narrower, but coefficients remain unbiased.. Then, when dropping values outside new OBW, the OBW became even smaller, and the estimated coefficients are biased downwards")`
